---
title: "Why does preregistration increase the persuasiveness of evidence?"
subtitle: "A Bayesian rationalization"
author: 
  - "Aaron Peikert"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
repro:
  packages:
    - xaringanthemer
    - showtext
    - svglite
    - patchwork
    - here
    - qrcode
---
class: center, middle
name: qrcode

```{r setup, include=FALSE}
library(xaringanthemer)
library(tidyverse)
library(patchwork)
library(qrcode)
options(htmltools.dir.version = FALSE)
knitr::read_chunk("R/plots.R")
knitr::opts_chunk$set(
  fig.width=9, fig.height=4, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE,
  dev = "svglite"
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
primary_color <- "#024959"
secondary_color <- "#F28705"
style_duo(
  primary_color = primary_color,
  secondary_color = secondary_color,
  header_font_google = google_font("Didact Gothic", "400", "400i"),
  text_font_google   = google_font("IBM Plex Sans", "400", "400i"),
  text_font_size = "1.5rem",
  )

style_extra_css(css = list(".todo" = list(color = "#17C25B"),
                           ".optional" = list(color = "#05BCE6"),
                           ".large" = list(`font-size` = "130%"),
                           ".small" = list(`font-size` = "80%"),
                           ".full" = list(padding = "0px",
                                          margin = "0px",
                                          border = "0px"),
                           ".vertical" = list(`writing-mode` = "vertical-rl",
                                              `text-orientation` = "mixed"),
                           ".blurry" = list(color = "transparent",
                                            `text-shadow` = "0 0 5px rgba(0,0,0,0.5)")),
                )
```

```{r, echo=FALSE}
link <- "github.com/aaronpeikert/bayes-prereg"
```

```{r, echo=FALSE, out.width = "30%"}
fs::dir_create("images")
generate_svg(qr_code(link), here::here("images/", "qr_slides.svg"), foreground = secondary_color, background = primary_color, show = FALSE)
knitr::include_graphics("images/qr_slides.svg")
```

`r link`

.pull-right[![CC0](https://mirrors.creativecommons.org/presskit/buttons/88x31/svg/cc-zero.svg)]

---
class: center, middle

# Why does preregistration increase the persuasiveness of evidence?

## A Bayesian rationalization

### Aaron Peikert

### DGPS | `r Sys.Date()`

---
class: inverse, center, middle

# Why do we preregister?

???

An objective for preregistration.
If commiting to specific research question and data analysis before gathering data is to be a useful tool, we must know to what end we employ it.

If you are uncertain about why you should preregister, I hope to provide some clarity.
If you think you know why you should preregister, I hope to sow some doubt.

---

```{r, echo=FALSE}
crisis <- tribble(
  ~response, ~bin, ~prop,
  "Yes, a signicant crisis", "YES", .52,
  "Yes, a slight crisis", "YES", .38,
  "No, there is no crisis", "NO", .03,
  "Don’t know", "NA", 0.07
)

crisis_plot <- crisis %>%
  group_by(bin) %>%
  summarise(prop = sum(prop)) %>%
  ungroup() %>% 
  ggplot(aes(
    "response",
    prop,
    fill = bin,
    label = bin,
    size = prop
  )) +
  geom_col() +
  geom_text(position = position_stack(vjust = 0.5)) +
  coord_flip() +
  theme_void() +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA)
  ) +
  scale_y_continuous(expand=c(0,0)) +
  scale_x_discrete(expand=c(0,0), limits = NULL) +
  scale_fill_manual(values = c("YES" = "#E3350E", "NO" = "#F28705", "NA" = "lightgrey")) +
  scale_size_continuous(range = c(4, 20)) +
  NULL
```

## Is there a replication crisis?*

.center[

```{r, out.width = '70%'}
crisis_plot /
  (crisis_plot +
     aes(label = str_c(prop * 100, "%")) + 
     scale_size_continuous(range = c(3, 12))) + 
  plot_layout(heights = c(.7, .3)) + 
  plot_annotation(theme = theme_void() + theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA)
))

```
]

.small.right[Baker, M. 1,500 scientists lift the lid on reproducibility. *Nature* 533, 452–454 (2016). https://doi.org/10.1038/533452a]

--

.small.right[***They call it "reproducibility" and fail to publish the raw data.**]

???

The simple answer is that we preregister because we think it is a remedy against the replication crisis.

---
class: center, middle
layout: true
---

## Because researchers are *fraudsters*?

--

# No

???

Some say to prevent intentional wrongdoing. 
However, preregistrations are easily circumvented with bad intentions.
We must assume researchers have good intentions.

---

## Because researchers are *fools*?

--

# No
--
, not necessarily

???

Some say to prevent unintentional wrongdoing.
Researchers must be put in restrains to prevent them from cutting themselves on the sharp edge of "false positives".

---
layout: false
class: inverse, center, middle

# Traditional view

---
class: center

# Researchers confuse:


.pull-left.center[

# confirmatory

]

.pull-right.center[

# exploratory

]

---
class: center

# Preregistration separates:

.pull-left.center[

# confirmatory

]

.pull-right.center[

# exploratory

]

--

.pull-left.center[

# ≈

# preregistered

]

.pull-right.center[


# ≈

# not preregistered

]

--

# ?

---
class: middle

.center[

## When is one preregistration better than another?

]

---
class: middle

.large[

What must be **included** in the preregistration?

When is a preregistration **detailed** enough?

Can one **deviate** from the preregistration?

]

---
class: inverse, center, middle

# A Formal Objective

---
class: middle

.large[

Ideally:

1. The original author has **good intentions**.
2. Their readers are **rational**.
3. The goal is to **convince** the readers.

]

---
layout: true
class: center, middle
---

.large[To connect rationality to probability imagine:]

---

## Researchers as gamblers

---

.large[

They bet on studies:

hoping to gain knowledge, and 

have to pay with hard work.

]

---

.left[The Dutch book argument:]

## Any rational better must follow the laws of probability.

--

.right[Together with well intentions, we may use Bayes˟.]

---

.large[

$$
\frac{P(H)P(E|H)}{P(E)}
$$

]

---

.large[

$$
\frac{\color{#F28705}{P(H)}P(E|H)}{P(E)}
$$

]

.right[The prior probability has nothing to do with preregistration.]

---

.large[

$$
\frac{P(H)\color{#F28705}{P(E|H)}}{P(E)}
$$

]

.right[The true positive rate / power has nothing to with preregistration.]


---

.large[

$$
\frac{P(H)P(E|H)}{\color{#F28705}{P(E)}}
$$

]

.right[The probability of evidence is the only thing left.]


---

.left[Law of total probability:]

.large[

$$
P(E) = P(H)P(E|H) + P(¬H)\color{#F28705}{P(E|¬H)}
$$

]

--

$P(H)$, $P(¬H)$, $P(E|H)$, are already ruled out.

---
## $P(E|¬H)$

The probability of finding evidence, even though the theory does not hold.

---
class: center, middle, inverse

# Theoretical Risk

---

Related concepts

---

## Theoretical risk > α-error

---

## Researchers' degrees of freedom > theoretical risk

---
class: center, middle, inverse

# Effect

---
name: theoretical-risk-plot

```{r}
results <- readr::read_rds(here::here("data", "simulation.rds"))
```

```{r}
results %>%
  filter(name == "Posterior Probability", certainty == "certain") %>%
  ggplot() +
  geom_line(aes(tau, value)) +
  xaringanthemer::theme_xaringan() +
  theme(
    legend.position = c(.2, .8),
    legend.title = element_blank(),
    text = element_text(size = 18, color = secondary_color),
    title = element_blank(),
    axis.title = element_text(size = 18),
    plot.caption = element_text(size = 15)
  ) +
  labs(x = "Theoretical Risk",
       y = "Posterior Probability",
       caption = "P(E|H) = .8, P(H) = .1") +
  NULL
```

---

### So maximize theoretical risk?

--

# No
--
, not necessarily

---

## Theories determine $P(E|¬H)$ (theoretical risk)

---

## Resources limit $P(E|H)$ (detectability)

---

## $P(E|H) \gg P(E|¬H)$

## detectability $\gg$ theoretical risk

*(statistical relevancy)*

---

Maximizing theoretical risk / minimizing researchers degrees of freedom is 

# unsustainable

--

Other ideas?

---
class: center, middle, inverse

# .blurry[Uncertainty]

---
template: theoretical-risk-plot

---
name: theoretical-risk-plot2

```{r measure-plots}
# see R/plots.R
```

```{r}
plots[["Posterior Probability"]] +
  xaringanthemer::theme_xaringan() +
    theme(legend.position = c(.2, .8),
        legend.title = element_blank(),
        text = element_text(size = 18, color = secondary_color),
        title = element_blank(),
        axis.title = element_text(size = 18),
        plot.caption = element_text(size = 15)) +
  labs(x = "Theoretical Risk",
       y = "Posterior Probability",
       caption = "P(E|H) = .8, P(H) = .1") +
  NULL
```

---

.large[**Preregistrations**] reduce .large[**uncertainty**] about .large[**theoretical risk**].

---

# Can I explore the data?

--

# Yes.

---
template: theoretical-risk-plot2

---

# Can I deviate?

--

# Sometimes.

---
template: theoretical-risk-plot2

---

# How much detail?

--

# Very detailed.

---
template: theoretical-risk-plot2

---
template: qrcode

---

### One further idea

---

# Preregistration as Code

### version control + dynamic documents

---
class: center

.pull-left[

### Standard Preregistration

hunches

↓

preregistration

↓

data

↓

article draft
]

.pull-right[
### Preregistration as Code

simulated data

↓

article draft with mock results

↓

data

↓

article draft with real results

]

---
template: qrcode
